### 2023-02-04 manuscript/manuscript.tex
1st commit

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  super,
  preprint,
  3p]{elsarticle}

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\journal{To be determined}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-num}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Extracorporeal CPR for Refractory Out-of-Hospital Cardiac Arrest},
  pdfauthor={James M Brophy},
  pdfkeywords={extracorporeal CPR, Bayesian statistics},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\setlength{\parindent}{6pt}
\begin{document}

\begin{frontmatter}
\title{Extracorporeal CPR for Refractory Out-of-Hospital Cardiac
Arrest \\\large{A Bayesian Perspective} }
\author[1]{James M Brophy%
\corref{cor1}%
\fnref{fn1}}
 \ead{james.brophy@mcgill.ca}

\affiliation[1]{organization={McGill University Health Center, Centre
for Health Outcomes Research (CORE)},addressline={5252 Boul. de
Maisonneuve West Room 2B.37},city={Montreal},postcode={H4A
3S5},postcodesep={}}

\cortext[cor1]{Corresponding author}
\fntext[fn1]{JMB is a research scholar supported by Les Fonds de
Recherche Québec Santé}

\begin{abstract}
A recent randomized clinical trial reported in patients with refractory
out-of-hospital cardiac arrest, extracorporeal CPR and conventional CPR
had similar effects on survival with a favorable neurologic outcome.
Herein, it is examined whether a Bayesian perspective allows any
additional insights into the interpretation of this trial.
\end{abstract}

\begin{keyword}
    extracorporeal CPR \sep 
    Bayesian statistics
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, borderline west={3pt}{0pt}{shadecolor}, breakable, frame hidden, enhanced, interior hidden, boxrule=0pt]}{\end{tcolorbox}}\fi

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Out-of-hospital cardiac arrest is a frequent event and fortunately its
devastating consequences can be partially mitigated by rapid
commencement of basic life support with high-quality chest compressions
and external defibrillation (conventional cardiopulmonary resuscitation
(CPR)). However, there remains a substantial subset of individuals who
do not respond rapidly to these measures and whether more invasive
measures. Whether the addition of more aggressive measure including
extracorporeal CPR (the addition of extracorporeal membrane oxygenation
to standard advanced cardiac life support) can improve survival and
diminish anoxic brain injury is a current topic of research. The largest
randomized clinical trial (RCT) examining this question recently
published their results\citep{CPR2023a}. For the primary outcome, 30 day
survival without significant neurological deficit, the authors observed
an odds ratio of 1.4 (95\% confidence interval, 0.5 to 3.5; P = 0.52) in
favor of extracorporeal CPR for leading to their conclusion ``In
patients with refractory out-of-hospital cardiac arrest, extracorporeal
CPR and conventional CPR had similar effects on survival with a
favorable neurological outcome''.\citep{CPR2023a}

This communication does not reiterate the many reasons to be wary of
null hypothesis significance testing (NHST), p values and confidence
intervals\citep{RN5420}. Rather it assumes the reader has perhaps heard
that Bayesian methods mirror our intuitive learning process and is
curious about its potential application to RCT interpretations.

Therefore the goal of this communication is to examine whether a
Bayesian perspective permits additional insights into the specific
clinical question regarding any added value of extracorporeal CPR
following an out-of-hospital arrest in patients refractory to standard
CPR.

\hypertarget{methods}{%
\section{Methods}\label{methods}}

The data for the primary outcome, 30 day survival with intact
neurological status, based on an intention to treat (ITT) analysis was
abstracted from the original INCEPTION trial \citep{CPR2023a} and used
for the primary analysis. The ITT analysis has the advantage of
minimizing bias by preserving the prognostic balance afforded by
randomization as well as assuring the validity of the statistical
analyses. ITT assesses subjects based on the group they were initially
(and randomly) allocated to, regardless of whether or not they dropped
out, were fully adhered to the treatment or switched to an alternative
treatment. ITT analyses can therefore be seen as a conservative estimate
which mirrors clinical effectiveness. In contrast, a per protocol (PP)
analysis involves a comparison of treatment groups in a trial that
includes only those patients who completed the treatment they were
originally allocated to. Similarly an ``as treated'' analysis considers
only which treatments subjects received, regardless of their
randomization status and protocol adherence. While both PP and
as-treated analyses alone may lead to bias, in conjunction with an ITT
analysis they may provide additional insights into efficacy and have
also been examined from a Bayesian perspective.

Bayesian analytical approaches provide a number of benefits over the
classical NHST approach, including parameter estimation accompanied by
direct probability statements about parameters of interest (herein the
risk of survival with intact neurological status), and the incorporation
prior knowledge \citep{BrophyCardio, Zampieri}.

These probability statements arise from the posterior distribution
according to the following equation:
\[ \text{Posterior}  = \frac{\text{Probability of the data} * \text{Prior}}{\text{Normalizing Constant}} \]

Therefore, in addition to the current data summarized by the probability
of the data (likelihood function) one requires a prior probability
distribution for each parameter. The mechanics of the Bayesian analyses
were performed using the Stan programming language \citep{stan} through
the R package rstanarm \citep{rstanarm} and fit a logistic regression
model with a single treatment parameter, \(\theta\). Because our focus
is the interpretation of the INCEPTION trial alone, our primary analysis
used rstanarm's default vague parameter priors
(\(log(\theta) \sim Normal [0, 2.50]\), thereby assuring that the
posterior distribution is dominated by the observed INCEPTION data.

The robustness of the Bayesian approach is often assessed by sensitivity
analyses that examine the variation in the posterior probability as a
function of the choice of different prior distributions. Using prior
information also underscores the important advantage of Bayesian
analyses to learn sequentially. There were two previous RCTs examining
extracorporeal CPR{[}\citet{RN6759}{]}\citep{RN6751} and while the
protocols are not identical, it may be reasonable to allow this data can
serve as an informed prior which can be updated with the INCEPTION data.
This prior information of the probability of success in each arm,
\(X_i\), can be summarized by a Beta distribution as follow
\[X_i\sim\mbox{Beta}(\alpha_0,\beta_0)\] where in the context of
Bernoulli trials, \(\alpha\) can be interpreted as 1 + number of
successes and \(\beta\) can be interpreted as 1 + number of failures.
The mean and variance of this distribution are
\[\mu=\frac{\alpha}{\alpha+\beta}\]

\[\sigma^2=\frac{\alpha\beta}{(\alpha+\beta)^2(\alpha+\beta+1)}\] where
\(\alpha\) and \(\beta\) are the number of successes and failures
respectively. The difference between the two treatment arms of these
previous studies can then serve as the prior information which in the
Bayesian paradigm can be updated with the INCEPTION data to form a
posterior distribution.

Posterior distributions are summarized with medians and 95\%
highest-density intervals (credible intervals), defined as the narrowest
interval containing 95\% of the probability density function
\citep{mcelreath2020}. We not only calculated the posterior probability
of any benefit (OR \textgreater1.00), but also of clinically meaningful
benefits (defined as OR \textgreater1.10).

All analyses were executed within the integrated development environment
of RStudio and the statistical code can be found on Github
(https://github.com/brophyj/eCPR).

\hypertarget{results}{%
\section{Results}\label{results}}

The ITT data from the three pertinent trials is shown in Table 1.
Performing a Bayesian analysis on the INCEPTION trial, using a vague
prior, produces an odds ratio 1.32 (95\% Credible Interval (CrI) 0.54 -
3.22). The closeness of this result to the original analysis confirms
the lack if impact of the vague prior and shows this Bayesian analysis
is dominated by the observed data.

\newpage

\hypertarget{tables}{%
\section{Tables}\label{tables}}

\begin{longtable}[]{@{}lrrrr@{}}
\caption{Table 1 Extracted ITT trial data}\tabularnewline
\toprule()
Trial & Fail CPR & Fail eCPR & Success CPR & Success eCPR \\
\midrule()
\endfirsthead
\toprule()
Trial & Fail CPR & Fail eCPR & Success CPR & Success eCPR \\
\midrule()
\endhead
INCEPTION & 52 & 56 & 10 & 14 \\
ARREST & 15 & 8 & 0 & 6 \\
PRAGUE & 108 & 86 & 24 & 38 \\
\bottomrule()
\end{longtable}

\hypertarget{bayesian-analysis}{%
\section{Bayesian analysis}\label{bayesian-analysis}}

\begin{verbatim}

Attaching package: 'LearnBayes'
\end{verbatim}

\begin{verbatim}
The following object is masked from 'package:rstanarm':

    laplace
\end{verbatim}

\begin{verbatim}

Model Info:
 function:     stan_glm
 family:       binomial [logit]
 formula:      prop_success ~ Tx
 algorithm:    sampling
 sample:       4000 (posterior sample size)
 priors:       see help('prior_summary')
 observations: 2
 predictors:   2

Estimates:
              mean    sd      10%     50%     90%  
(Intercept) -1.5987  0.2265 -1.8906 -1.5963 -1.3088
TxeCPR       0.1542  0.0501  0.0897  0.1535  0.2181

Fit Diagnostics:
           mean    sd      10%     50%     90%  
mean_PPD 11.9740  3.1650  8.0000 12.0000 16.0000

The mean_ppd is the sample average posterior predictive distribution of the outcome variable (for details see help('summary.stanreg')).

MCMC diagnostics
              mcse   Rhat   n_eff
(Intercept)   0.0042 0.9995 2872 
TxeCPR        0.0009 0.9999 2827 
mean_PPD      0.0544 0.9994 3387 
log-posterior 0.0238 1.0006 1835

For each parameter, mcse is Monte Carlo standard error, n_eff is a crude measure of effective sample size, and Rhat is the potential scale reduction factor on split chains (at convergence Rhat=1).
\end{verbatim}

\textbf{Have obviously not done this correctly as the prior should have
shifted the default model parameter for eCPR to a larger number with a
small sd}

\hypertarget{figures}{%
\section{Figures}\label{figures}}

\begin{verbatim}
Warning: Removed 9 rows containing missing values (`stat_slabinterval()`).
\end{verbatim}

\begin{verbatim}
Warning: Using the `size` aesthietic with geom_segment was deprecated in ggplot2 3.4.0.
i Please use the `linewidth` aesthetic instead.
\end{verbatim}

\includegraphics{manuscript_files/figure-pdf/unnamed-chunk-4-1.pdf}

\includegraphics{manuscript_files/figure-pdf/unnamed-chunk-4-2.pdf}

\newpage

\renewcommand\refname{References}
  \bibliography{bibliography.bib}

\end{document}