### 2023-02-11 manuscript/manuscript.qmd
starting discussion

---
title: A Bayesian Perspective 
subtitle: Extracorporeal CPR for Refractory Out-of-Hospital Cardiac Arrest
author:
  - name: James M Brophy MD PhD
    email: james.brophy@mcgill.ca
    affiliations: 
        - id: McGill University 
          name: McGill University Health Center
          department: Centre for Health Outcomes Research (CORE) 
          address: 5252 Boul. de Maisonneuve West Room 2B.37
          city: Montreal
          state: QC
          postal-code: H4A 3S5
    attributes:
        corresponding: true
    note: JMB is a research scholar supported by Les Fonds de Recherche Québec Santé
abstract: |
  A recent randomized clinical trial, INCEPTION, reported in patients with refractory out-of-hospital cardiac arrest, extracorporeal CPR (eCPR) and conventional CPR (cCPR) had similar effects on survival with a favorable neurologic outcome. The current study examines if a Bayesian perspective provides  additional insights. Depending on the prior selected, the Bayesian approach for the INCEPTION intention-to treat (ITT) analysis shows an equivalence probability between 13.4 - 16.8% (defined as 1 / 1.1 < odds ratio (OR) < 1.1). The probability of clinical superiority with eCPR ranges from 65.7 - 77.0 % (defined as OR > 1.1). A similar analyses using INCEPTION per protocol (PP) data shows an equivalence probability between 4.7 - 20.2% with reduced probabilities of clinical superiority not exceeding 25%. It is concluded that a Bayesian perspective allows considerable additional insights into the trial analysis and interpretation. The non-negligible probabilities of increased survival with their considerable residual uncertainties, suggests that additional studies are required before concluding that eCPR and cCPR have similar survival effects.
keywords: 
  - extracorporeal CPR
  - Bayesian statistics
date: last-modified
bibliography: bibliography.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: To be determined
      formatting: preprint
      model: 3p
      cite-style: super
editor: 
  markdown: 
    wrap: 72
---

```{r echo=FALSE,  message=FALSE, warning=FALSE, include=FALSE, results='hide'}
knitr::opts_chunk$set(echo = FALSE)
options(scipen = 1, digits = 3)
pander::panderOptions('keep.trailing.zeros', TRUE)

library(here)
library(tidyverse)
library(broom)
library(cobalt)
library(table1)
library(knitr)
library(patchwork)
library(rstanarm)
library(tidybayes)
library(gt)

```

# Introduction

Out-of-hospital cardiac arrest is a frequent event and fortunately its
devastating consequences can be partially mitigated by rapid
commencement of basic life support with high-quality chest compressions
and external defibrillation (conventional cardiopulmonary resuscitation
(cCPR)). However, there remains a substantial subset of individuals who
do not respond rapidly to these measures and whether more invasive
measures. Whether the addition of more aggressive measure including
extracorporeal CPR (the addition of extracorporeal membrane oxygenation
to standard advanced cardiac life support (eCPR)) can improve survival and
diminish anoxic brain injury is a current topic of research. A large
randomized clinical trial (RCT) examining this question recently
published their results[@CPR2023a]. For their primary outcome, 30 day
survival without significant neurological deficit, an odds ratio of 1.4 (95% confidence interval, 0.5 to 3.5; P = 0.52) in
favor of extracorporeal CPR was observed leading to the conclusion "In patients
with refractory out-of-hospital cardiac arrest, extracorporeal CPR and
conventional CPR had similar effects on survival with a favorable
neurological outcome".[@CPR2023a]

This communication does not reiterate the many reasons to be wary of
null hypothesis significance testing (NHST), p values and confidence
intervals[@RN5420]. Rather it assumes the reader has perhaps heard that
Bayesian methods mirror our intuitive learning and diagnostic processes and is curious
about its potential application to RCT analyses and interpretations.

Therefore the goal of this communication is to examine whether a
Bayesian perspective permits additional insights into the specific
clinical question regarding any added value of eCPR
following an out-of-hospital arrest in patients refractory to cCPR.

# Methods

The data for the primary outcome, 30 day survival with intact
neurological status, based on an intention to treat (ITT) analysis was
abstracted from the original INCEPTION trial [@CPR2023a] and used for
the primary analysis. The ITT analysis has the advantage of minimizing
bias by preserving the prognostic balance afforded by randomization as
well as assuring the validity of the accompanying statistical analyses.

Bayesian analytical approaches provide a number of benefits over the
classical NHST approach, including parameter estimation accompanied by
direct probability statements about parameters of interest (herein the
risk of survival with intact neurological status), and the incorporation
prior knowledge [@BrophyCardio; @Zampieri].

These probability statements arise from the posterior distribution
according to the Bayes Thereom, expressed as follows: 
$$ \text{Posterior}  = \frac{\text{Probability of the data} * \text{Prior}}{\text{Normalizing Constant}} $$

Therefore, in addition to the current data summarized by the probability
of the data (likelihood function), prior probability
distributions are required. Because our main focus is the analysis and 
interpretation of the INCEPTION trial alone, our primary analysis used a default vague parameter prior
($log(\theta) \sim Normal [0, 2.50]$, thereby assuring that the
posterior distribution is dominated by the observed INCEPTION data.

The robustness of the Bayesian approach is often assessed by sensitivity
analyses that examine the variation in the posterior probability as a
function of the choice of different prior distributions. Incorporating prior
information underscores another important advantage of Bayesian
analyses, the ability to learn sequentially. There were two previous RCTs examining
extracorporeal CPR[@RN6759; @RN6751] and while the protocols are not
identical, it may be reasonable to allow this data to serve as 
informed priors for the eCPR parameter, which can be updated with the INCEPTION data.

Therefore, in addition to the vague prior described above, we considered three possible informative priors    
i) a **combined** prior using all the available prior RCT data[@RN6751;@RN6759]               
ii) an **enthusiastic** prior, so labelled since this uses only the ARREST data, a trial stopped prematurely for efficacy        
iii) a **skeptical** prior, so labelled since this uses only the PRAGUE data, a trial  stopped prematurely for futility

This prior
information of the probability of eCPR success in each previous trial, $X_i$, can be summarized as a normal distribution with a mean equal to the proportion of successes, $\hat{p_i}$ with a standard deviation equal to $$\sqrt{\hat{p_i}*(1-\hat{p_i})}$$ 
As  baseline success rates for cCPR varies markedly between the three studies, it was decided to maintain the INCEPTION control baseline with the vague prior for all analyses and to update only the eCPR arm with prior information.

Posterior distributions are summarized with medians and 95%
highest-density intervals (credible intervals (CrI)), defined as the narrowest
interval containing 95% of the probability density function
[@mcelreath2020]. Bayesian analyses permit not only calculations of the posterior probability of
any additional survival with eCPR (OR \>1.00), but also of clinically meaningful benefits. While there is no universal definition for a clinically meaningful benefit, a survival OR \>1.10 may be an acceptable threshold for many. Bayesian analyses also allows calculation of the probability between any two points. For example, rather than simply comparing if the survival of one treatment is better than another, one can calculate a range of practical equivalence (ROPE) between treatments. While different ranges may be proposed, +/- 10% seems a reasonable small difference that many woul consider as equivalent.

ITT assesses subjects based on the group they were initially (and randomly) allocated
to, regardless of whether or not they dropped out, were fully adhered to
the treatment or switched to an alternative treatment. In superiority trials, ITT analyses can
therefore be seen as a conservative estimate which mirrors clinical
effectiveness. In contrast, a per protocol (PP) analysis accounts for adherence by analyzing only those
patients who completed the treatment they were originally allocated to.
While PP may provide additional insights
into efficacy, it is subject to bias. Since a PP analysis when performed in conjunction with an ITT analysis may provide additional insights, this has also been subjected to a Bayesian analysis.

Posterior distributions were estimated using the Hamiltonian Monte Carlo,a form of 
Markov Chain Monte Carlo simulations in which the gradient of the log posterior is 
used to efficiently sample the posterior space. This was implemented in `Stan`[@stan] using the front end `rstanarm` package[@rstanarm] by fitting a logistic regression model with a single treatment parameter. All analyses were executed using `R`[@R]  within the integrated development environmentof RStudio[@Rstudio].  The statistical code can be found on Github
(https://github.com/brophyj/eCPR).

# Results
```{r cache=TRUE}
df <- tibble(Trial = c("INCEPTION", "INCEPTION","ARREST", "ARREST", "PRAGUE", "PRAGUE"), Tx = c("CPR", "eCPR", "CPR", "eCPR", "CPR", "eCPR"),
             fail = c(52, 56, 15, 8, 108, 86),
             success = c(10,14, 0, 6,24,38 )) %>% 
  mutate(total = fail + success,
         prop_success = success / total) 
df1 <- df[, c(1:4)]  %>% tidyr::pivot_wider(names_from = Tx, values_from = c(fail,success))
#frequentist INCEPTION  to reproduce NEJM result
fit <- glm(prop_success ~ Tx, data = df[c(1:2),], family = binomial(link="logit"), weights = total)
# summary(fit, digits=3)
# print(c(exp(.262),exp(.262+c(-1,1)*1.96*.457)))
# [1] 1.300 0.531 3.183

# - INCEPTION alone default vague priors
set.seed(1234)
fit1 <- stan_glm(prop_success ~ Tx, data = df[c(1:2),], family = binomial(link="logit"), weights = total, refresh=0) #stan_glm with combined prior
fit1_sum <- summary(fit1, digits=3)
conf_95 <- c(exp(fit1_sum[2]),exp(fit1_sum[2]+c(-1,1)*1.96*fit1_sum[10]))
tt <- fit1 %>%
    spread_draws(`(Intercept)`, `TxeCPR`)
fig_prob <- sum(tt$TxeCPR>0)/nrow(tt)
# summary(fit1, digits=3)
# print(c(exp(.2785),exp(.2785+c(-1,1)*1.96*.4537)))
# get_variables(fit1)
# prior_summary(fit1)
# help('prior_summary.stanreg') 
# P > 1.1
fig_prob_1.1 <- sum(tt$TxeCPR > 0.0953)/nrow(tt) # OR >1.1 (log (1.1) = 0.0953)
# [1] 0.657

# P > 1/1.1
fig_prob_1.1.1 <- sum(tt$TxeCPR > -0.0954)/nrow(tt) # OR > 1 / 1.1 (log (0.909) = -0.0954) = 0.791

# difference equivalence

vague_eq <- fig_prob_1.1.1 - fig_prob_1.1
```

ITT data from the INCEPTION trial[@CPR2023a] and two other pertinent trials[@RN6759;@RN6751] that also randomized out of hospital cardiac arrest patients to cCPR to eCPR are shown in Table 1. Performing a Bayesian analysis on the INCEPTION trial, using a default vague prior, produces an odds ratio (OR) 1.32 with 95% CrI 0.54 - 3.22. The closeness of this result to the original analysis (OR, 1.4; 95% CI 0.5 - 3.5) confirms the minimal impact of the default vague prior and reveals a Bayesian analysis completely dominated by the observed INCEPTION data.

One of the advantages of a Bayesian approach is the ability to make direct probability statements about the estimand of improved eCPR survival. The eCPR probability density function for improved survival from INCEPTION data with the default vague prior is displayed in Figure 1 and reveals that the probability of enhanced survival with eCPR is `r round(fig_prob*100,1)`%. The probability that the improved survival exceeds a 10% improvement is `r round(fig_prob_1.1*100,1)`% and the ROPE probability is `r round(vague_eq*100,1)`% (Table 2).

```{r priors, cache=TRUE}
## running using prior for eCPR from different studies but leaving CPR with vague prior and INCEPTION data since baseline risks with CPR seem 
## to vary widely btw the different studies

eCRP_bayes_no_intercept_prior <- function(data, m_eCPR, sd_eCPR ) {
  set.seed(1234)
  fit <- stan_glm(prop_success ~ Tx, data = data, family = binomial(link="logit"), weights = total, refresh=0,
                  prior = normal(m_eCPR, sd_eCPR))
  fit_sum <- summary(fit, digits=3)
  conf_95 <- c(exp(fit_sum[2]),exp(fit_sum[2]+c(-1,1)*1.96*fit_sum[10]))
  tt <- fit %>%
    spread_draws(`(Intercept)`, `TxeCPR`)
  fig_prob <- sum(tt$TxeCPR>0)/nrow(tt)
  fit_draws <- fit %>% tidybayes::tidy_draws() 
  my_list <- list(fig_prob, conf_95, fit_draws)
  return(my_list)
}

# combined_prior 
# informative priors using 2 previous studies combined
comb_m_CPR <- sum(df1[c(2,3),4]) / (sum(df1[c(2,3),4]) + sum(df1[c(2,3),2])) # proportion previous results for CPR
comb_sd_CPR <- sqrt(comb_m_CPR*(1-comb_m_CPR))                               
comb_m_eCPR <- sum(df1[c(2,3),5]) / (sum(df1[c(2,3),5]) + sum(df1[c(2,3),3])) # proportion previous results for eCPR
comb_sd_eCPR <- sqrt(comb_m_eCPR*(1-comb_m_eCPR)) #sd

# enthusiastic prior with only ARREST study
enth_m_CPR <- sum(df1[c(2),4]) + 1 / (sum(df1[c(2),4]) + sum(df1[c(2),2])) # proportion previous results for CPR, + 0.5 to avoid 0 cell
enth_sd_CPR <- sqrt(enth_m_CPR*(1-enth_m_CPR))
enth_m_eCPR <- sum(df1[c(2),5]) / (sum(df1[c(2),5]) + sum(df1[c(2),3])) # proportion previous results for eCPR
enth_sd_eCPR <- sqrt(enth_m_eCPR*(1-enth_m_eCPR)) # sd

# skeptical prior with only PRAGUE study
skep_m_CPR <- sum(df1[c(3),4])  / (sum(df1[c(3),4]) + sum(df1[c(3),2])) # proportion previous results for CPR
skep_sd_CPR <- sqrt(skep_m_CPR*(1-skep_m_CPR))
skep_m_eCPR <- sum(df1[c(3),5]) / (sum(df1[c(3),5]) + sum(df1[c(3),3])) # proportion previous results for eCPR
skep_sd_eCPR <- sqrt(skep_m_eCPR*(1-skep_m_eCPR)) #sd

comb_prior1 <-eCRP_bayes_no_intercept_prior(df[c(1:2),], comb_m_eCPR, comb_sd_eCPR) #stan_glm with combined prior
enth_prior1 <- eCRP_bayes_no_intercept_prior(df[c(1:2),], enth_m_eCPR, enth_sd_eCPR) #stan_glm with enthusiastic prior
skep_prior1 <- eCRP_bayes_no_intercept_prior(df[c(1:2),], skep_m_eCPR, skep_sd_eCPR) #stan_glm with skeptical prior

# >OR 1.1
comb_prob_1.1 <- sum(comb_prior1[[3]]$TxeCPR > 0.0953) / nrow(comb_prior1[[3]]) # OR >1.1 (log (1.1) = 0.0953) = 0.732
enth_prob_1.1 <- sum(enth_prior1[[3]]$TxeCPR > 0.0953) / nrow(enth_prior1[[3]]) # OR >1.1 (log (1.1) = 0.0953) = 0.77
skep_prob_1.1 <- sum(skep_prior1[[3]]$TxeCPR > 0.0953) / nrow(skep_prior1[[3]]) # OR >1.1 (log (1.1) = 0.0953) = 0.71

# > OR 1/ 1.1 (0.909)
comb_prob_1.1.1 <- sum(comb_prior1[[3]]$TxeCPR > -0.0954) / nrow(comb_prior1[[3]]) # OR > 1 / 1.1 (log (0.909) = -0.0954) = 0.884
enth_prob_1.1.1 <- sum(enth_prior1[[3]]$TxeCPR > -0.0954) / nrow(enth_prior1[[3]]) # OR >1 / 1.1 (log (0.909) = -0.0954) = 0.908
skep_prob_1.1.1 <- sum(skep_prior1[[3]]$TxeCPR > -0.0954) / nrow(skep_prior1[[3]]) # OR >1 / 1.1 (log (0.909) = -0.0954) = 0.878

# equivalence +/- 10%
comb_eq <- comb_prob_1.1.1 - comb_prob_1.1 
enth_eq <- enth_prob_1.1.1 - enth_prob_1.1
skep_eq <- skep_prob_1.1.1 - skep_prob_1.1
```

Three different informative priors were considered i)  a N(`r round(comb_m_eCPR,2)`, `r round(comb_sd_eCPR,2)`) ii) a N(`r round(enth_m_eCPR,2)`, `r round(enth_sd_eCPR,2)`) iii) a N(`r round(skep_m_eCPR,2)`, `r round(skep_sd_eCPR,2)`) distributions to represent all the combined available RCT data[@RN6759;@RN6751], only the ARREST[@RN6751] data  and only the PRAGUE[@RN6759] data, respectively

These different prior probabilities were updated using with the INCEPTION ITT data to create the posterior distributions displayed in Table 2. The probability for enhanced eCPR survival has increased to 80.4% with a skeptical prior to 84.9% with the enthusiastic prior and, as expected, the associated uncertainty has been reduced, as reflected by the narrower 95% CrI, with the additional data. The posterior probability of improved eCPR survival goes from 80% with a skeptical prior to 85% with the enthusiastic prior. 
The probability that the eCPR survival improvements exceed the proposed 10% clinical threshold are `r round(skep_prob_1.1*100,1)`% and `r round(enth_prob_1.1*100,1)`% for the skeptical and enthusiastic priors, respectively. The corresponding ROPE probabilities are `r round(skep_eq*100,1)`%  and `r round(enth_eq*100,1)`% (Table 2). Posterior probabilities with the combined prior fell as expected between the results with the skeptical and enthusiastic priors. The graphical presentations of these results are shown in Figure 2.

INCEPTION did not report a per-protocol analysis. Such an analysis may be helpful in assessing treatment efficacy, unfortunately at the risk of an increased risk of bias by not respecting the ITT principle. From INCEPTION Figure S4[@CPR2023a], it appears that the per protocol data for mortality is 13 survivors from 61 patients in the cCPR group compared to 5 survivors from 46 patients receiving eCPR. With a vague prior, the OR of increased survival with eCPR is decreased compared to cCPR but with very wide CrI (OR 0.45, 95% CrI 0.15 - 1.35), limiting any definitive conclusions. The decreased eCPR success rates in the INCEPTION PP data results in reduced posterior probabilities of eCPR benefit and an increased probability of equivalence or benefit with cCPR. The exact probabilities of increased eCPR survival incorporating the previous identified informative priors are shown in Table 3.

```{r PP, cache=TRUE}
# per protocol - only data for INCEPTION, other data just placeholder if PP data becomes available
df_pp <- tibble(Trial = c("INCEPTION", "INCEPTION","ARREST", "ARREST", "PRAGUE", "PRAGUE"), Tx = c("CPR", "eCPR", "CPR", "eCPR", "CPR", "eCPR"),
             fail = c(48, 41, 15, 8, 108, 86),
             success = c(13,5, 0, 6,24,38 )) %>% 
  mutate(total = fail + success,
         prop_success = success / total)

# df_pp[c(1:2),] data set used for PP analyses

set.seed(1234)
fit1_pp <- stan_glm(prop_success ~ Tx, data = df_pp[c(1:2),], family = binomial(link="logit"), weights = total, refresh=0) #stan_glm with combined prior
fit1_pp_sum <- summary(fit1_pp, digits=3)
conf_95_pp <- c(exp(fit1_pp_sum[2]),exp(fit1_pp_sum[2]+c(-1,1)*1.96*fit1_pp_sum[10]))
tt_pp <- fit1_pp %>%
    spread_draws(`(Intercept)`, `TxeCPR`)
fig_prob_pp <- sum(tt_pp$TxeCPR>0)/nrow(tt_pp)
# summary(fit1, digits=3)
# print(c(exp(.2785),exp(.2785+c(-1,1)*1.96*.4537)))
# get_variables(fit1)
# prior_summary(fit1)
# help('prior_summary.stanreg') 
fig_prob_1.1_pp <- sum(tt_pp$TxeCPR>0.0953)/nrow(tt_pp) #
fig_prob_1.1.1_pp <- sum(tt_pp$TxeCPR>-0.0954)/nrow(tt_pp)

comb_prior1_pp <-eCRP_bayes_no_intercept_prior(df_pp[c(1:2),], comb_m_eCPR, comb_sd_eCPR) #stan_glm with combined prior
enth_prior1_pp <- eCRP_bayes_no_intercept_prior(df_pp[c(1:2),], enth_m_eCPR, enth_sd_eCPR) #stan_glm with enthusiastic prior
skep_prior1_pp <- eCRP_bayes_no_intercept_prior(df_pp[c(1:2),], skep_m_eCPR, skep_sd_eCPR) #stan_glm with skeptical prior

# >OR 1.1
comb_prob_pp_1.1 <- sum(comb_prior1_pp[[3]]$TxeCPR > 0.0953) / nrow(comb_prior1_pp[[3]]) # OR >1.1 (log (1.1) = 0.0953) = 0.732
enth_prob_pp_1.1 <- sum(enth_prior1_pp[[3]]$TxeCPR > 0.0953) / nrow(enth_prior1_pp[[3]]) # OR >1.1 (log (1.1) = 0.0953) = 0.77
skep_prob_pp_1.1 <- sum(skep_prior1_pp[[3]]$TxeCPR > 0.0953) / nrow(skep_prior1_pp[[3]]) # OR >1.1 (log (1.1) = 0.0953) = 0.71

# > OR 1/ 1.1 (0.909)
comb_prob_pp_1.1.1 <- sum(comb_prior1_pp[[3]]$TxeCPR > -0.0954) / nrow(comb_prior1_pp[[3]]) # OR > 1 / 1.1 (log (0.909) = -0.0954) = 0.884
enth_prob_pp_1.1.1 <- sum(enth_prior1_pp[[3]]$TxeCPR > -0.0954) / nrow(enth_prior1_pp[[3]]) # OR >1 / 1.1 (log (0.909) = -0.0954) = 0.908
skep_prob_pp_1.1.1 <- sum(skep_prior1_pp[[3]]$TxeCPR > -0.0954) / nrow(skep_prior1_pp[[3]]) # OR >1 / 1.1 (log (0.909) = -0.0954) = 0.878

# equivalence +/- 10%
vague_eq_pp <- fig_prob_1.1.1_pp - fig_prob_1.1_pp
comb_eq_pp <- comb_prob_pp_1.1.1 - comb_prob_pp_1.1
enth_eq_pp <- enth_prob_pp_1.1.1 - enth_prob_pp_1.1
skep_eq_pp <- skep_prob_pp_1.1.1 - skep_prob_pp_1.1

```

# Discussion

This Bayesian analysis of the INCEPTION ITT data alone (with a vague prior) suggests the presence of a non-negligible probability of benefit with eCPR, in the range of 73%, compared to cCPR. The probability that this improvement in survival exceeds 10% is almost 66%. The consideration of pertinent prior information from two previous RCTs[@RN6751;@RN6759] in the field results in estimates of improved eCPR survival probabilities, with intact neurological status, in the vicinity of 80-85%. Moreover depending on the choice of prior beliefs included, the posterior probability of at least a 10% improved eCPR survival is in the range of 71 - 77% probability. In contrast to the INCEPTION original conclusion of similar survival effects between the two treatment, this re-analysis suggests only a modest probability of approximately that survival probabilities between the two techniques are within a reasonably acceptable equivalence range. The overall message would appear to be that while any definitive conclusions regarding the superiority of eCPR are lacking, the possibility of a clinically meaningful benefit or even less likley clinically meaningful harm has not been reasonably excluded and continued research is necessary to clarifying the remaining uncertainties.

The INCEPTION researchers are to be congratulated on their trial design, its execution, and addressing an important clinical question in the most challenging of research environments. However the natural constraints of standard statistical analyses, also known as null hypothesis significance testing, limits the full exploitation of the current data and updating of past knowledge. For trails failing to meet statistical significance, those incorrectly labelled "negative" trials, null hypothesis significance testing favors confusion "between absence of evidence and evidence of absence"[@RN6765].

\newpage

# Tables

```{r}
# knitr::kable(df1, caption = "Table 1 Extracted ITT trial data", col.names=c("Trial", "Fail CPR", "Fail eCPR", "Success CPR", "Success eCPR"))

gt(df1) %>%
  tab_header(
    title = "Table 1 Extracted ITT trial data"
  ) %>%
  cols_label(
    Trial = html("Trial"),
    fail_CPR = html("Fail CPR (n)"),
    fail_eCPR = html("Fail eCPR (n)"),
    success_CPR = html("Success CPR (n)"),
    success_eCPR = html("Success eCPR (n)"),
  ) %>%
  tab_style(style = cell_text(align = 'center'), locations = cells_column_labels()) %>%
  cols_align(
  align = c("center"),
  columns = c(fail_CPR, fail_eCPR, success_CPR, success_eCPR )
  ) %>%
  tab_source_note(
    source_note = "eCPR = extracorporeal cardiopulmonary resuscitation"
  )

## Table 2

table_tbl_2 <- 
  tibble(
    prior = c("Vague", "Combined", "Enthusiastic", "Skeptical"),
    pt_est = c(conf_95[1], comb_prior1[[2]][1], enth_prior1[[2]][1], skep_prior1[[2]][1]),
    llimit = c(conf_95[2], comb_prior1[[2]][2], enth_prior1[[2]][2], skep_prior1[[2]][2]),
    ulimit = c(conf_95[3], comb_prior1[[2]][3], enth_prior1[[2]][3], skep_prior1[[2]][3]),
    p.gt.1 = c(fig_prob, comb_prior1[1], enth_prior1[1], skep_prior1[1]),
    p.gt.1.1 = c(fig_prob_1.1, as.numeric(comb_prob_1.1), as.numeric(enth_prob_1.1), as.numeric(skep_prob_1.1)),
    p.eq   = c(vague_eq, comb_eq, enth_eq, skep_eq)
  )

gt(table_tbl_2) %>%
  tab_header(
    title = "Table 2 eCPR odds ratios, 95% credible intervals and probabilities with various priors"
  ) %>%
  tab_source_note(
    source_note = "Vague: default vague prior"
  ) %>%
  tab_source_note(
    source_note = "Combined: prior eCPR data from ARREST + PRAGUE"
  ) %>%
  tab_source_note(
    source_note = md("Enthusiastic: prior eCPR data from ARREST alone")
  ) %>%
  tab_source_note(
    source_note = md("Skeptical: prior eCPR data from PRAGUE alone")
  ) %>%
  tab_source_note(
    source_note = md("ROPE: range of practical equivalence = + / - 10% OR (odds ratio)")
  ) %>%
  tab_spanner(
    label = "Priors",
    columns = c(prior)
   ) %>%
  tab_spanner(
    label = "OR",
    columns = c(pt_est)
  ) %>%
  tab_spanner(
    label = "Probabilities",
    columns = c(p.gt.1, p.gt.1.1, p.eq)
   ) %>%
  tab_spanner(
    label = "95% CrI",
    columns = c(llimit, ulimit)
  ) %>%
  cols_label(
    prior = html(""),
    pt_est = html("point estimate"),
    llimit = html("lower limit"),
    ulimit = html("upper limit"),
    p.gt.1 = html("p(OR) >1 "),
    p.gt.1.1 = html("p(OR) >1.1 "),
    p.eq   = html(" p(ROPE)")
  ) %>%
  tab_style(style = cell_text(align = 'center'), locations = cells_column_labels()) %>%
  cols_align(
  align = c("center"),
  columns = c(pt_est, llimit, ulimit, p.gt.1.1, p.eq) 
  ) %>%
  fmt_number(
    columns = c(pt_est, llimit, ulimit, p.gt.1.1),
    decimals = 3,
    drop_trailing_zeros = FALSE
)

## Table 3

table_tbl_3 <- 
  tibble(
    prior = c("Vague", "Combined", "Enthusiastic", "Skeptical"),
    pt_est = c(conf_95_pp[1], comb_prior1_pp[[2]][1], enth_prior1_pp[[2]][1], skep_prior1_pp[[2]][1]),
    llimit = c(conf_95_pp[2], comb_prior1_pp[[2]][2], enth_prior1_pp[[2]][2], skep_prior1_pp[[2]][2]),
    ulimit = c(conf_95_pp[3], comb_prior1_pp[[2]][3], enth_prior1_pp[[2]][3], skep_prior1_pp[[2]][3]),
    p.gt.1 = c(fig_prob_pp, as.numeric(comb_prior1_pp[1]), as.numeric(enth_prior1_pp[1]), as.numeric(skep_prior1_pp[1])),
    p.gt.1.1 = c(fig_prob_1.1_pp, as.numeric(comb_prob_pp_1.1), as.numeric(enth_prob_pp_1.1), as.numeric(skep_prob_pp_1.1)),
    p.eq   = c(vague_eq_pp, comb_eq_pp, enth_eq_pp, skep_eq_pp)
  )

gt(table_tbl_3) %>%
  tab_header(
    title = "Table 3 eCPR (per protocol) odds ratios, 95% credible intervals and probabilities with various priors"
  ) %>%
  tab_source_note(
    source_note = "Vague: default vague prior"
  ) %>%
  tab_source_note(
    source_note = "Combined: prior eCPR data from ARREST + PRAGUE"
  ) %>%
  tab_source_note(
    source_note = md("Enthusiastic: prior eCPR data from ARREST alone")
  ) %>%
  tab_source_note(
    source_note = md("Skeptical: prior eCPR data from PRAGUE alone")
  ) %>%
  tab_source_note(
    source_note = md("ROPE: range of practical equivalence = + / - 10% OR (odds ratio)")
  ) %>%
  tab_spanner(
    label = "Priors",
    columns = c(prior)
   ) %>%
  tab_spanner(
    label = "OR",
    columns = c(pt_est)
  ) %>%
  tab_spanner(
    label = "Probabilities",
    columns = c(p.gt.1, p.gt.1.1, p.eq)
   ) %>%
  tab_spanner(
    label = "95% CrI",
    columns = c(llimit, ulimit)
  ) %>%
  cols_label(
    prior = html(""),
    pt_est = html("point estimate"),
    llimit = html("lower limit"),
    ulimit = html("upper limit"),
    p.gt.1 = html("p(OR) >1 "),
    p.gt.1.1 = html("p(OR) >1.1 "),
    p.eq   = html(" p(ROPE)")
  ) %>%
  tab_style(style = cell_text(align = 'center'), locations = cells_column_labels()) %>%
  cols_align(
  align = c("center"),
  columns = c(pt_est, llimit, ulimit, p.gt.1, p.gt.1.1, p.eq)
  ) %>%
  fmt_number(
    columns = c(pt_est, llimit, ulimit, p.gt.1, p.gt.1.1, p.eq) ,
    decimals = 3,
    drop_trailing_zeros = FALSE
)

```

\newpage

# Figures

```{r fig1, warning=FALSE, message=FALSE, fig.height=6}

fit1 %>% tidybayes::tidy_draws() |> 
  ggplot(aes(x = exp(TxeCPR),
             fill = after_stat(x > 1),
             slab_alpha = after_stat((x)))
  ) +
  ggdist::stat_halfeye() +
  scale_x_continuous(name="OR survival with extracorporeal CPR", limits=c(0, 5)) +
  scale_y_continuous(name="Density") +
  labs(title = "Figure 1 INCEPTION ITT analysis with vague prior", caption = "Clinical superiority = AUC to right of vertical line at HR = 1.1 = 65.7%\nthin horizontal line at 0 = 95% CrI with black circle = point estimate\nthick horizontal line = +/- 68% CrI", hjust=0) +
  annotate(geom="text", x=2.0,y=.6, label = "Probability of increased survival \n(with intact neurological status) \nwith extracorporeal CPR = 72.7% \n(light turquoise area under the curve (AUC) \nto the right of HR=1)", hjust =0) +
  geom_segment(aes(x = 1.1, y = 0, xend = 1.1, yend = .90)) +
  geom_segment(aes(x = 1 / 1.1, y = 0, xend = 1 / 1.1, yend = .86)) +
  annotate(geom="text", x=2.0,y=.35, label = "Probability (range) of practical equivalence (+/- 10% OR) \nAUC between black vertical lines = 13.4% ", hjust = 0) +
  theme_classic() +
  theme(legend.position="none")

```

\newpage

```{r fig2, warning=FALSE, message=FALSE, fig.height=18, fig.width=8}
g1 <- comb_prior1[[3]]  %>% tidybayes::tidy_draws() |> 
  ggplot(aes(x = exp(TxeCPR),
             fill = after_stat(x > 1),
             slab_alpha = after_stat((x)))
  ) +
  ggdist::stat_halfeye() +
  scale_x_continuous(name="OR survival with extracorporeal CPR", limits=c(0, 5)) +
  scale_y_continuous(name="Density") +
  labs(title = "Figure 2a INCEPTION ITT analysis with combined prior*", subtitle = "*data from ARREST and PRAGUE",
       caption = "Clinical superiority = AUC to right of vertical line at HR = 1.1 = 73.2%") +
  annotate(geom="text", x=2.0,y=.6, label = "Probability of increased survival \n(with intact neurological status) \nwith extracorporeal CPR = 81.2% \n(light turquoise area under the curve (AUC) \nto the right of HR=1)", hjust = 0) +
  geom_segment(aes(x = 1.1, y = 0, xend = 1.1, yend = .86)) +
  theme_classic() +
  theme(legend.position="none")

g2 <- enth_prior1[[3]]  %>% tidybayes::tidy_draws() |> 
  ggplot(aes(x = exp(TxeCPR),
             fill = after_stat(x > 1),
             slab_alpha = after_stat((x)))
  ) +
  ggdist::stat_halfeye() +
  scale_x_continuous(name="OR survival with extracorporeal CPR", limits=c(0, 5)) +
  scale_y_continuous(name="Density") +
  labs(title = "Figure 2b INCEPTION ITT analysis with enthusiastic prior*", subtitle = "*data from ARREST alone",
       caption = "Clinical superiority = AUC to right of vertical line at HR = 1.1 = 77.0%") +
  annotate(geom="text", x=2.0,y=.6, label = "Probability of increased survival \n(with intact neurological status) \nwith extracorporeal CPR = 84.9% \n(light turquoise area under the curve (AUC) \nto the right of HR=1)", hjust = 0) +
  geom_segment(aes(x = 1.1, y = 0, xend = 1.1, yend = .77)) +
  theme_classic() +
  theme(legend.position="none")

g3 <- skep_prior1[[3]]  %>% tidybayes::tidy_draws() |> 
  ggplot(aes(x = exp(TxeCPR),
             fill = after_stat(x > 1),
             slab_alpha = after_stat((x)))
  ) +
  ggdist::stat_halfeye() +
  scale_x_continuous(name="OR survival with extracorporeal CPR", limits=c(0, 5)) +
  scale_y_continuous(name="Density") +
  labs(title = "Figure 2c INCEPTION ITT analysis with skeptical prior*", subtitle = "*data from PRAGUE alone",
       caption = "Clinical superiority = AUC to right of vertical line at HR = 1.1 = 71.0%") +
  annotate(geom="text", x=2.0,y=.6, label = "Probability of increased survival \n(with intact neurological status) \nwith extracorporeal CPR = 80.4% \n(light turquoise area under the curve (AUC) \nto the right of HR=1)", hjust = 0) +
  geom_segment(aes(x = 1.1, y = 0, xend = 1.1, yend = .88)) +
  theme_classic() +
  theme(legend.position="none")

g1 / g2 / g3 +
  plot_annotation(title = 'Figure 2. Probability density plots with informative priors')

```

\newpage

# References {.unnumbered}